---
title: "Getting started"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(collapse = T, comment = "#>")
NOT_CRAN <- identical(tolower(Sys.getenv("NOT_CRAN")), "true")
knitr::opts_chunk$set(purl = NOT_CRAN)
```

```{r}
library(rscorecard)
```

Using `{rscorecard}` to download data from the College Scorecard API
requires two steps:

1. Setting your API key
2. Making a request

## 1. Setting your API key

If you don't already have one, reqest your (free) API key from
[https://api.data.gov/signup](https://api.data.gov/signup). It should
only take a few moments to register and receive your key.

Once you've gotten your key, you can store it usig `sc_key()`. In the
absence of a key value argument, `sc_get()` will search your R
environment for `DATAGOV_API_KEY`. It will complete the data request
if found. `sc_key()` command will store your key in `DATAGOV_API_KEY`,
which will persist until the R session is closed.

```{r, eval = FALSE}
# NB: You must use a real key, of course... 
sc_key('xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx')
```

If you want a more permanent solution, you can add the following line
(with your actual key, of course) to your `.Renviron` file.  See this
[appendix](ftp://cran.r-project.org/pub/R/web/packages/httr/vignettes/api-packages.html)
for more information.

```{r, eval = FALSE}
# NB: You must use a real key, of course... 
DATAGOV_API_KEY=xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx
```

## Simple request

Each request requires the following four commands piped together using
`%>%`:

1. `sc_init()`
2. `sc_filter()`
3. `sc_select()`
4. `sc_get()`

The command chain must begin with `sc_init()` and end with
`sc_get`. All other commands can come in any order.

The request belower should return a tibble with the name, IPEDS ID,
state, and degree-seeking undergrad enrollment of all primarily
Baccalaureate colleges in the Mid East region located in rural areas:

```{r}
df <- sc_init() %>% 
    sc_filter(region == 2, ccbasic == c(21,22,23), locale == 41:43) %>% 
    sc_select(unitid, instnm, stabbr, ugds) %>% 
    sc_get()
```
```{r}
df
```

Because we didn't include a specific year, the `latest` data are
returned. We could have specifically asked for the latest data using
`sc_year('latest')`:

```{r}
df <- sc_init() %>% 
    sc_filter(region == 2, ccbasic == c(21,22,23), locale == 41:43) %>% 
    sc_select(unitid, instnm, stabbr, ugds) %>%
    sc_year('latest') %>% 
    sc_get()
```
```{r}
df
```

For a prior year's data, change the value in `sc_year()`:


```{r}
df <- sc_init() %>% 
    sc_filter(region == 2, ccbasic == c(21,22,23), locale == 41:43) %>% 
    sc_select(unitid, instnm, stabbr, ugds) %>%
    sc_year(2005) %>% 
    sc_get()
```
```{r}
df
```

### Field of study data

In the fall of 2019, the College Scorecard released field of
study-level data elements (4 digit CIP code level). These data
elements can be requested alongside institution-level data:

```{r}
df <- sc_init() %>% 
    sc_filter(region == 2, ccbasic == c(21,22,23), locale == 41:43) %>% 
    sc_select(unitid, instnm, stabbr, ugds, cipcode, cipdesc, debt_mdn) %>%
    sc_year("latest") %>% 
    sc_get()
```

```{r}
## filter to show only those with non-NA values for median debt
df %>% dplyr::filter(!is.na(debt_mdn))
```

**Important note:**

The mapping scheme of data across years isn't consistent across data
elements. From the [technical documentation for institution-level
data](https://collegescorecard.ed.gov/assets/FullDataDocumentation.pdf):

> The data contain diverse measures of institutional performance
> constructed both with an eye towards the type of information that
> would be most useful to prospective students, as well as towards how
> the measures might promote accountability for institutions. The
> measures require different definitions of cohorts. Users of the data
> should be aware of this, particularly when constructing analyses of
> the relationship between different measures. Moreover, reporting
> inaccuracies in some data elements used for cohort definitions are
> also important. (p. 37)

That is, while the _reporting_ year (_e.g._, `sc_year(2016)`) may be
the same, the _measurement_ year may not directly align. The same
holds true when trying to align institution-level data with field of
study-level data (see the [technical documentation for field of
study-level
data](https://collegescorecard.ed.gov/assets/FieldOfStudyDataDocumentation.pdf)
for more information). 

The upshot is that `{rscorecard}` will return data based on what the
API call returns, but the user should take care to ensure that
returned data elements align with expectations and project needs.

# More information and examples

For more information about each command, see
[Commands](https://www.btskinner.io/rscorecard/articles/commands.html).

For more examples, see [More
examples](https://www.btskinner.io/rscorecard/articles/more_examples.html).








